|

### 2023-05-03 paper.Rmd
fix some minor issues

We can assess residuals in multiple ways. To examine the their univariate distribution, residuals may be plotted as a histogram or normal probability plot. Using the classical normal linear regression model as an example, if the distribution is symmetric and unimodal, we would consider it to be well-behaved. However, if the distribution is skewed, bimodal, multimodal, or contains outliers, there would be cause for concern. We can also inspect the distribution by conducting a goodness-of-fit test, such as the Shapiro-Wilk normality test [@shapiro1965analysis].

In addition, scatterplots of residuals plotted against the fitted values and each of the explanatory variables, can be used to scrutinize their relationships. If there are any visually discoverable patterns, the model is potentially inadequate or incorrectly specified. In general, one looks for noticeable departures from the model such as non-linear pattern or heteroskedasticity. A non-linear pattern would suggest that the model needs to have some additional non-linear terms. Heteroskedasticity suggests that the error is dependent on the predictors, and hence violates the independence assumption. However, correctly judging whether NO pattern exists in a residual plot is a difficult task for humans. Humans will almost always will see a pattern (see @kahneman2011thinking), so the question that really needs answering is whether any pattern perceived is consistent with randomness, purely sampling variability, or noise. It is especially difficult to teach this to new analysts and students [@loy2021bringing]. To address this, statistical tests are available to check for non-linear patterns [e.g. @ramsey1969tests], and heteroskedasticity  [e.g. @breusch1979simple].

@@

Visual patterns of non-linearity are simulated using four different orders of probabilist's Hermite polynomials ($j = 2, 3, 6, 18$). The values of $j$ were chosen so that distinct shapes of non-linearity were included in the residual plot. These include "U", "S", "M" and "triple-U" shape as shown in Figure \ref{fig:different-shape-of-herimite}. A greater value of $j$ will result in a curve with more turning points. It is expected that the "U" shape will be the easiest detect, as the shape gets more complex it will be harder to perceive in a scatterplot, particularly when there is noise. Figure \ref{fig:different-sigma} shows the "U" shape for different amounts of noise ($\sigma$).

>>>>

We can assess residuals in multiple ways. To examine the univariate distribution, residuals may be plotted as a histogram or normal probability plot. Using the classical normal linear regression model as an example, if the distribution is symmetric and unimodal, we would consider it to be well-behaved. However, if the distribution is skewed, bimodal, multimodal, or contains outliers, there would be cause for concern. We can also inspect the distribution by conducting a goodness-of-fit test, such as the Shapiro-Wilk normality test [@shapiro1965analysis].

In addition, scatterplots of residuals plotted against the fitted values and each of the explanatory variables, can be used to scrutinize their relationships. If there are any visually discoverable patterns, the model is potentially inadequate or incorrectly specified. In general, one looks for noticeable departures from the model such as non-linear pattern or heteroskedasticity. A non-linear pattern would suggest that the model needs to have some additional non-linear terms. Heteroskedasticity suggests that the error is dependent on the predictors, and hence violates the independence assumption. However, correctly judging whether NO pattern exists in a residual plot is a difficult task for humans. Humans will almost always see a pattern (see @kahneman2011thinking), so the question that really needs answering is whether any pattern perceived is consistent with randomness, purely sampling variability, or noise. It is especially difficult to teach this to new analysts and students [@loy2021bringing]. To address this, statistical tests are available to check for non-linear patterns [e.g. @ramsey1969tests], and heteroskedasticity  [e.g. @breusch1979simple].

@@

Visual patterns of non-linearity are simulated using four different orders of probabilist's Hermite polynomials ($j = 2, 3, 6, 18$). The values of $j$ were chosen so that distinct shapes of non-linearity were included in the residual plot. These include "U", "S", "M" and "triple-U" shape as shown in Figure \ref{fig:different-shape-of-herimite}. A greater value of $j$ will result in a curve with more turning points. It is expected that the "U" shape will be the easiest to detect, as the shape gets more complex it will be harder to perceive in a scatterplot, particularly when there is noise. Figure \ref{fig:different-sigma} shows the "U" shape for different amounts of noise ($\sigma$).
